---
title: "Circular Drift Difussion Model on JAGS: Full example"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1.25cm,right=1.25cm,top=1.25cm,bottom=1.25cm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Generate/load simulated data

\vspace{-0.2in}

```{r, warning=FALSE, fig.width=7,fig.height=4, fig.align='center'}
# Establish no. of trials
trials <- 350
# Call Rscript to generate simulated data / load it  if already existing
source("./getData.R")
dim(data)
# Plot data
cddm.plotData(data)
```

\vspace{-0.5in}

```{r}
# Print parameter values used to generate this data
par
```

\clearpage

# 2. Write JAGS model

```{r}
modelFile <- "cddm.bug"
write('
            model{
                  # Likelihood
                    for (i in 1:N) {
                         X[1:2,i] ~ dcddm(drift, bound, ter0, theta0)
                    }
                      
                  # Priors
                    drift  ~ dnorm(0, 1)
                    theta0 ~ dnorm(0, 1)T(-3.14, 3.14)
                    bound  ~ dgamma(3, 2)
                    ter0   ~ dexp(1)T(, tmin)
                  }',
      modelFile)
```


where: 

- `drift` is the magnitude of the drift vector composed by the individual drift rates related to the average motion observed across the x and y axes, according to the CDDM.
- `bound` is the threshold (i.e. the radius of the circle)
- `ter0` is the non-decision time (a.k.a. "time for encoding and response")
- `theta0` is the direction of the drift vector, in radians.

# Prepare Settings to be passed to JAGS

```{r}
n.chains = 4
n.iter = 2500
n.burnin = 500
n.thin = 1
perParticipant = FALSE
perTask = FALSE

sampling.Settings <- list(n.chains,n.iter,n.burnin,n.thin,perParticipant,perTask)
names(sampling.Settings) <- c("n.chains","n.iter","n.burnin","n.thin","perParticipant","perTask")
```

# Run JAGS model

```{r}
source("../Functions/runCDDMjags.R")

samplesFile <- "samples.RData"

if(file.exists(samplesFile)){
   load(file=samplesFile)
}else{
   myJAGSsampling.CDDM(sampling.Settings,modelFile,samplesFile,data)}
```

\clearpage

```{r}
source("../Functions/plotJAGSsamples.R")
plot.ShowAllChains(samples)
```

```{r}
source("../Functions/processJAGSsamples.R")
myJAGSsampling.Rhat.max(samples)
```